Truthfulness Verification System

### 2011-04-27 trunk/proposal/paper.tex
EI NAO MUSA, ar commit commit koro na :D
updated with totha's alg

\section {Proposed System}

\begin{figure}
  \centering
  \subfloat[System
  overview]{\label{fig:System}\includegraphics[width=0.85\textwidth]{system.pdf}}
  \caption{System overview}
  \label{fig:system}
\end{figure}

\subsection{Description of extraction algorithm}

@@

\section{Use of Wikipedia}
\label{sec:UseOfWikipedia}
We have tried two approaches for using the Wikipedia. In both we utilize the content words to identify articles and there upon treat them separately. In the first approach we treat the articles as bag of words and do not consider the sentence structure and in the next approach we retain the logical structure of each sentence and try to infer from word overlaps.
%When we do a regular expression search of the $DU$ "Les Paul" in these two sentences, we do not get any match. Hence we can conclude that the statement is not true.

\subsection{Bag of words}
\begin{enumerate}
\item Step I: Detection of wiki articles from $DU$, $AU$, $TU$.
\begin {enumerate}
\item For the $AU$s, we take the words as they appear in the output of T-verifier(mainly because most of them are single word). For $TU$ we use a basic exponential algorithm that takes all possible combinations of the content words.
\item The default configuration of T-verifier generates a total of 5 $AU$s, without any ranking.  For each content word in a $AU$, or a combination of words in case the $AU$ consists of multiple content words, we check if there exists a wikipedia article. There may be three cases:
\begin{enumerate}
\item No article. Continue with next word or word combination. 
\item One article. Go to next step
\item More than one article. The page lands into a "disambiguation" page. We discuss this later. 
\end{enumerate}
\end{enumerate}
\item We remove the html tags from the pages, and convert them into simple bag of words. Repetation of a word if any is retained.
\item We select the following pairwise combination - a wikiarticle detected from $AU$, ($W_{au_{i}}$) and a wikiarticle detected from $TU$ ($W_{tu_{j}}$).
\item We take the common words $c_{ij} = {w_{1}, w_{2}, \ldots}$ matched by regular expressions in these two bags.
\item Since this is a regular expression matching we will have two different values for $c_{ij}$ = common($W_{au_{i}}$, $W_{tu_{j}}$) and $c_{ji}$ = common($W_{tu_{j}}$, $W_{au_{i}}$). This is explained below.
\item In this collection of common words, $c_{ij}$, we search each of the content words of the $TU$, not included in the title of $W_{tu_{j}}$ 
\item We observe the maximum number of matches in all of such interesctions and use the count to rank the $AU$-s generated by T-verifer. 
\end{enumerate}

\paragraph{Handling Disambiguation}
For the testing that we have done with the five problematic sentences, we manually first used the disambiguation page itself in the above process. But as expected, it did not yeild good results. We are experimenting with the following ideas for disabmigutation.
\begin{enumerate}
\item Apply the above steps for each of the page linked from the disambiguation page and get the one which gives the maximum count
\item Use the heading under which it appears in the disambiguation page, the one sentence summary appearing adjacent to the entry in the disambiguation page, and the "Categories" under which a particular article has been tagged.
\end{enumerate}
For this report, we manually checked the disambiguation pages by searching them for content words. This significantly improved the results.

\paragraph{Word matching}
We found that the intersection of the two bags of words, $W_{au_{i}}$ and $W_{tu_{i}}$ produced different results based on the ordereing of the articles while performing the intersection operation. This is because the regular expression match finds the maximal match. Lets consider that the first file, $W_{au_{i}}$ has "invent", and the second file, $W_{tu_{i}}$ has "invented", and "inventor". So when we pick a word in $W_{au_{i}}$ and try to match words in $W_{tu_{i}}$, we will find 2 matches. However, the words from second file, both "invented", and "inventor" will not find any matches in ($W_{au_{i}}$). Taking both way intersections will ensure we do not miss out different morphological variations of the same word appearing in the context common to the both articles.

\subsection{Examples}
Examples of the algorithm described above is discussed below. The algorithm 2 produce perfect result for three of the five sentences. But it struggles with the remaining two sentences. However, the algorithm 2 primarily augment the algorithm 1 and add strength to the results from algorithm 1. Hence they work better together and produce overall better output.

\subsubsection{Tom Hanks was the lead actress  int move 'Sleepless in Seattle'}
The alternative units for this sentence are "Tom Hanks", "Meg Ryan", "Hanks Meg", "Nora", and "Sam". Here the sentence containing "Tom Hanks" contain the word actor but not actress. So the algorithm classify it as false. But the sentence containing "Meg Ryan" has the word actress and hence classify it as true. For other alternative units, we don't even get any sentences containing the alternative unit. So the algorithm produce an unambiguous single true statement for this sentence.

\subsubsection{Les Paul invented the electric guitar}
The alternative units for this sentence are "Les Paul", "rickenbacker", "Gibson", "Leo", and "Lloyd". Now there is inherent ambiguity about the true inventor of the electric guitar. On reading the articles related above, there is no information that Rickenbacker actually inventing the electric guitar. He was a founding member of the Rickenbacker company that produces electric guitar. Les Paul on the other hand was a pioneer in designing and developing of what a prototype. Moreover, some of the categories attached with the Les Paul page are American musical instrument makers, Guitar makers, Inventors of musical instruments, National Inventors Hall of Fame inductees. In comparison Adolph Rickenbaker page has only one guitar related category - Guitar stubs, meaning the article is stuck in content.

\subsubsection{Luna 2 is the first spacecraft on the moon}
The alternative units for this sentence are "Luna 2", "Apollo", "Land", "Landing", and "Nasa". Our algorithm returns true for both Luna 2 and Apollo and return false for all others. Now, both Luna 2 and Apollo are in reality true considering the facts that Luna 2 was the first unmanned spacecraft and Apollo was the first manned spacecraft. Hence, we can't say one of the Luna 2 or Apollo as false if farther information is not given. So algorithm 2 can produce correct result for this sentence also.

\subsubsection{English is the Primary Language of the Philippines}
The alternative units for this sentence are "English", "Filipino", "Spanish", "Education", and "History". Our current algorithm struggle with this sentence because of occurrence of content words and alternative units in the same sentence for all the alternative units. So it classify all of them as true and assign same weight to all of them. However, as our algorithm 1 can rank the alternative units in correct order, merged rank from algorithm 1 and algorithm 2 is still correct and produce overall correct result.

\subsubsection{Sullivan is the fastest swimmer in the world}
The alternative units for this sentence are "Sullivan", "Michael phelps", "ALexander", "Long", and "Ernest". Similar to the previous sentence, the algorithm 2 can not produce good ranking for for this sentence. However, our algorithm 1 produces correct ranking and together they produce overall correct result.

\section{Conclusion}
This ranking procedure bolsters the result in the previous algorithm always without any exception. On observing the five examples above, we expect a combination of the two algorithms will give us a proper ranking of the alternate units for proper truth verification.

>>>>

\section{Proposed System}
\label{sec:UseOfWikipedia}
We have tried two approaches for using the Wikipedia. In both we utilize the content words to identify articles and there upon treat them separately. In the first approach we treat the articles as bag of words and do not consider the sentence structure and in the next approach we retain the logical structure of each sentence and try to infer from word overlaps.
%When we do a regular expression search of the $DU$ "Les Paul" in these two sentences, we do not get any match. Hence we can conclude that the statement is not true.

\subsection{Bag of words}

%\begin{algorithm}
%\small
%\caption{Ranking Algorithm 1}
%\label{alg:AssignScore1} 
%\begin{algorithmic}
%\STATE \textbf{Description}: Given the alternate units generated by the current T-verifier system, rank the units in descending order of truthfulness using Wikipedia
%\STATE \textbf{Input}: alter\_units.txt genereated by T-verifier. Each line in this file is a tuple containing the sentence id, the alternate units generated and the topic unit.
%\STATE \textbf{Output}: Ranked alter\_units 
%\FOR {each sentence $s_{i}$}
%	\STATE $tu_{i}$ = topic unit string
%	\STATE $s_{i}.tu\_list$ = find\_wikiarticles\_tu($tu_{s_{i}}$)
%		
%	\FOR {each alternate unit $s_{i}.au_{j}$ in $s_{i}.au\_list$ }
%		\STATE $bow_{s_{i}.au_{j}}$ = find\_wikiarticles\_au($s_{i}.au_{j}$)
%	\ENDFOR
%	\FOR {each alternate unit $s_{i}.au_{j}$ in $s_{i}.au\_list$ }
%		\STATE $bow_{s_{i}.au_{j}}$ = find\_wikiarticles\_au($s_{i}.au_{j}$)
%	\ENDFOR	
%	%\FOR{every wiki\_article $w_{p}$ in $list_wiki_articles_au_{s_{i}}$ }
%	%	$bow_w_{p}$ = generate\_bag\_of\_words($w_{p}$)	
%	%	\FOR {every wiki\_article $w_{j}$ in $list_wiki_articles_tu_{j}$  }	
%	%	\ENDFOR
%	%\ENDFOR
%\ENDFOR
%
%
%\end{algorithmic}
%\end{algorithm}
%

\begin{algorithm}
\small
\caption{Ranking Algorithm 1}
\label{alg:AssignScore1}
\begin{algorithmic}
\STATE \textbf{Description}: Given the alternate units generated by
the current T-verifier system, rank the units in descending order of
truthfulness using Wikipedia
\STATE \textbf{Input}: alter\_units.txt genereated by T-verifier. Each
line in this file is a tuple containing the sentence id, the alternate
units generated and the topic unit.
\STATE \textbf{Output}: Ranked alter\_units
\FOR {each sentence $s_{i}$}
       \STATE $tu_{s_{i}}$ = topic unit string
       \STATE $s_{i}.au\_list$ = alternate units generated by T-verifier
       \STATE $s_{i}.tu\_list$ = find\_wikiarticles\_tu($tu_{s_{i}}$)

       \FOR {each alternate unit $s_{i}.au_{j}$ in $s_{i}.au\_list$ }
               \STATE $w_{s_{i}.au_{j}}$ = find\_wikiarticles\_au($s_{i}.au_{j}$)
               \IF {$w_{s_{i}.au_{j}}$ == $\phi$}
                       \STATE skip to processing next $w_{s_{i}.au_{j}}$
               \ENDIF
               \IF {$w_{s_{i}.au_{j}}$ is diambiguation page}
                       \STATE $w_{s_{i}.au_{j}}$ = disambiguate($w_{s_{i}.au_{j}}$)
                       \STATE $bow_{s_{i}.au_{j}}$ = generate\_bag\_of\_words($s_{i}.au_{j}$)
               \ENDIF
       \ENDFOR
       \FOR {each $w_{s_{i}.tu_{j}}$ in $s_{i}.tu\_list$ }
               \STATE $bow_{s_{i}.tu_{j}}$ = generate\_bag\_of\_words($s_{i}.tu_{j}$)
       \ENDFOR

       \FOR {$s_{i}.au_{j}$ in $s_{i}.au\_list$}
               \FOR {each $w_{s_{i}.tu_{k}}$ in $s_{i}.tu\_list$ }
                       \STATE $ common\_words^{au_{j}}_{tu_{k}} $ =  $bow_{s_{i}.au_{j}}
\cap bow_{s_{i}.tu_{k}}$
                       \STATE $ remaining\_words_{s_{i}.tu_{k}}$  = $ tu_{s_{i}}$ - words
forming title of $w_{s_{i}.tu_{j}} $ - stop words
                       \FOR {each word $w$ in $ remaining\_words_{s_{i}.tu_{k}}$}
                               \STATE $s_{i}.au_{j}.score$ += count of $w$
                       \ENDFOR
               \ENDFOR
       \ENDFOR
\STATE print $s_{i}.au_{j}$ reverse sorted by $s_{i}.au_{j}.score$
\ENDFOR

\pagebreak

\begin{algorithm}
\small
\caption{Find wiki articles}
\label{alg:findwikiarticles}
\begin{algorithmic}
\STATE \textbf{Description}: Given a string of words determine phrases
that may be titles of wikipedia articles
\STATE \textbf{Input}: a string of words $s$
\STATE $l$ = length($s$)
\STATE $result$ = $\phi$
\FOR{$i$ in 1 to $l$}
       \FOR{$j$ in $i$ to $l$}
               \STATE $sub\_string_{ij}$ += $s_i$ + $blank$
       \ENDFOR
       \IF{ $sub\_string_{ij}$ consits only stopwords}
               \STATE discard and continue to next interation
       \ELSE
               \STATE search wikipedia with $sub_string_{ij}$
               \STATE $result_{sub\_string_{ij}}$ = (displaytitle, url, categories,
redirects) for $sub\_string_{ij}$
       \ENDIF
\ENDFOR
\RETURN result
\end{algorithmic}
\end{algorithm}

\pagebreak

\begin{algorithm}
\small
\caption{Disambiguate articles}
\label{alg:disambiguate}
\begin{algorithmic}
\STATE \textbf{Description}: Given a disambiguation page returns the
wikipedia article that is most relevant to setence being processed
\STATE \textbf{Input}: disambiguation page $p$,$tu_{s_{i}}$
\STATE extract all outlinks urls from the body of the page
\STATE form a dictionary $d$ using disambiguation element ($de$) and
the disambiguation description ($dd$) appearing adjacent to the $de$
\FOR {every $de$ in $d$}
       \STATE $de.ddscore$ = count content word overlap between $dd$ and
($tu_{s_{i}}$ - words in  $de$)
\ENDFOR
\STATE form a dictionary $l$ using disambiguation element ($de$) and
the lead section ($ls$) appearing as the first paragraph of the
article in $de$
\FOR {every $de$ in $l$}
       \STATE $de.lsscore$ = count content word overlap between $ls$ and
($tu_{s_{i}}$ - words in  $de$)
\ENDFOR
\STATE form a dictionary $c$ using disambiguation element ($de$) and
the category list ($cl$) appearing attached with the article on $de$
\FOR {every $de$ in $c$}
       \FOR {every category in $cl$}
               \STATE $de.clscore$ = count content word overlap between $cl$ and
($tu_{s_{i}}$ - words in  $de$)
       \ENDFOR
\ENDFOR
\STATE Reverse sort the $de$s in each method using the scores
\STATE Use Bordas ranking or weighted Bordas to merge the ranks
\RETURN top ranked $de$
\end{algorithmic}
\end{algorithm}

%\begin{enumerate}
%\item Step I: Detection of wiki articles from $DU$, $AU$, $TU$.
%\begin {enumerate}
%\item For the $AU$s, we take the words as they appear in the output of T-verifier(mainly because most of them are single word). For $TU$ we use a basic exponential algorithm that takes all possible combinations of the content words.
%\item The default configuration of T-verifier generates a total of 5 $AU$s, without any ranking.  For each content word in a $AU$, or a combination of words in case the $AU$ consists of multiple content words, we check if there exists a wikipedia article. There may be three cases:
%\begin{enumerate}
%\item No article. Continue with next word or word combination. 
%\item One article. Go to next step
%\item More than one article. The page lands into a "disambiguation" page. We discuss this later. 
%\end{enumerate}
%\end{enumerate}
%\item We remove the html tags from the pages, and convert them into simple bag of words. Repetation of a word if any is retained.
%\item We select the following pairwise combination - a wikiarticle detected from $AU$, ($W_{au_{i}}$) and a wikiarticle detected from $TU$ ($W_{tu_{j}}$).
%\item We take the common words $c_{ij} = {w_{1}, w_{2}, \ldots}$ matched by regular expressions in these two bags.
%\item Since this is a regular expression matching we will have two different values for $c_{ij}$ = common($W_{au_{i}}$, $W_{tu_{j}}$) and $c_{ji}$ = common($W_{tu_{j}}$, $W_{au_{i}}$). This is explained below.
%\item In this collection of common words, $c_{ij}$, we search each of the content words of the $TU$, not included in the title of $W_{tu_{j}}$ 
%\item We observe the maximum number of matches in all of such interesctions and use the count to rank the $AU$-s generated by T-verifer. 
%\end{enumerate}
%
%
%\paragraph{Handling Disambiguation}
%For the testing that we have done with the five problematic sentences, we manually first used the disambiguation page itself in the above process. But as expected, it did not yeild good results. We are experimenting with the following ideas for disabmigutation.
%\begin{enumerate}
%\item Apply the above steps for each of the page linked from the disambiguation page and get the one which gives the maximum count
%\item Use the heading under which it appears in the disambiguation page, the one sentence summary appearing adjacent to the entry in the disambiguation page, and the "Categories" under which a particular article has been tagged.
%\end{enumerate}
%For this report, we manually checked the disambiguation pages by searching them for content words. This significantly improved the results.
%
%\paragraph{Word matching}
%We found that the intersection of the two bags of words, $W_{au_{i}}$ and $W_{tu_{i}}$ produced different results based on the ordereing of the articles while performing the intersection operation. This is because the regular expression match finds the maximal match. Lets consider that the first file, $W_{au_{i}}$ has "invent", and the second file, $W_{tu_{i}}$ has "invented", and "inventor". So when we pick a word in $W_{au_{i}}$ and try to match words in $W_{tu_{i}}$, we will find 2 matches. However, the words from second file, both "invented", and "inventor" will not find any matches in ($W_{au_{i}}$). Taking both way intersections will ensure we do not miss out different morphological variations of the same word appearing in the context common to the both articles.

\paragraph{Examples}
Examples of the algorithm described above is discussed below. The algorithm 2 produce perfect result for three of the five sentences. But it struggles with the remaining two sentences. However, the algorithm 2 primarily augment the algorithm 1 and add strength to the results from algorithm 1. Hence they work better together and produce overall better output.

\paragraph{Tom Hanks was the lead actress  int move 'Sleepless in Seattle'}
The alternative units for this sentence are "Tom Hanks", "Meg Ryan", "Hanks Meg", "Nora", and "Sam". Here the sentence containing "Tom Hanks" contain the word actor but not actress. So the algorithm classify it as false. But the sentence containing "Meg Ryan" has the word actress and hence classify it as true. For other alternative units, we don't even get any sentences containing the alternative unit. So the algorithm produce an unambiguous single true statement for this sentence.

\paragraph{Les Paul invented the electric guitar}
The alternative units for this sentence are "Les Paul", "rickenbacker", "Gibson", "Leo", and "Lloyd". Now there is inherent ambiguity about the true inventor of the electric guitar. On reading the articles related above, there is no information that Rickenbacker actually inventing the electric guitar. He was a founding member of the Rickenbacker company that produces electric guitar. Les Paul on the other hand was a pioneer in designing and developing of what a prototype. Moreover, some of the categories attached with the Les Paul page are American musical instrument makers, Guitar makers, Inventors of musical instruments, National Inventors Hall of Fame inductees. In comparison Adolph Rickenbaker page has only one guitar related category - Guitar stubs, meaning the article is stuck in content.

\paragraph{Luna 2 is the first spacecraft on the moon}
The alternative units for this sentence are "Luna 2", "Apollo", "Land", "Landing", and "Nasa". Our algorithm returns true for both Luna 2 and Apollo and return false for all others. Now, both Luna 2 and Apollo are in reality true considering the facts that Luna 2 was the first unmanned spacecraft and Apollo was the first manned spacecraft. Hence, we can't say one of the Luna 2 or Apollo as false if farther information is not given. So algorithm 2 can produce correct result for this sentence also.

\paragraph{English is the Primary Language of the Philippines}
The alternative units for this sentence are "English", "Filipino", "Spanish", "Education", and "History". Our current algorithm struggle with this sentence because of occurrence of content words and alternative units in the same sentence for all the alternative units. So it classify all of them as true and assign same weight to all of them. However, as our algorithm 1 can rank the alternative units in correct order, merged rank from algorithm 1 and algorithm 2 is still correct and produce overall correct result.

\paragraph{Sullivan is the fastest swimmer in the world}
The alternative units for this sentence are "Sullivan", "Michael phelps", "ALexander", "Long", and "Ernest". Similar to the previous sentence, the algorithm 2 can not produce good ranking for for this sentence. However, our algorithm 1 produces correct ranking and together they produce overall correct result.

\subsection{Rank merging}
This ranking procedure bolsters the result in the previous algorithm always without any exception. On observing the five examples above, we expect a combination of the two algorithms will give us a proper ranking of the alternate units for proper truth verification.

\section {Alternative methods that are explored}

%\begin{figure}
%  \centering
%  \subfloat[System
%  overview]{\label{fig:System}\includegraphics[width=0.85\textwidth]{system.pdf}}
%  \caption{System overview}
%  \label{fig:system}
%\end{figure}
%
%
%\subsection{Description of extraction algorithm}

@@
